A Horrocks-type theorem for even orthogonal $K_2$

### 2020-04-06 dl-paper.tex
Restructuring...

For the rest of this subsection $\Phi$ is a simply-laced root system of rank $\geq 3$ and $\alpha$ is a fixed root of $\Phi$. Notice that in the simply-laced case the assumption $\alpha+\beta\not\in \Phi$ in the definition of $Z_0(\alpha)$ is superfluous, i.\,e. $Z_0(\alpha) = \{ \alpha\in\Phi \mid \alpha \perp \beta \}$
(cf.~\cite[Proposition~5.7]{St71}).

@@

We need one more technical definition. We denote by $Z_{\alpha, \beta}$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by elements $z_\gamma(f, X\xi)$, where $\gamma \in Z_+(\alpha) \setminus \{ \alpha - \beta \}$.

\begin{lemma} \label{image-K-a-b} The image of $K(\alpha, \beta)$ under conjugation by $x_\alpha(aX^{-1})$ is contained in the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by $K(\alpha, \beta)$ and $Z_{\alpha, \beta}$. \end{lemma}
\begin{proof} To simplify notation we call the generators of the subgroup $Z_{\alpha,\beta}$ ``generators of type Z``. By~\eqref{eq:Zconj1}, \eqref{eq:Zconj2} the conjugate of a generator $z_\gamma(Xf, \xi)$ of type 1 by $x_\alpha(aX^{-1})$ is either the generator itself or a product of generators of type Z, Z, 1. \cref{Crels} also shows that the conjugate of a generator of type 5 is either the generator itself or a product of generators of type Z, Z, 5.

The conjugation by $x_\alpha(aX^{-1})$ fixes both types of generators of type 3 and transforms $x_{-\beta}(X^2f)$ into a product of generators of type 2, 4. The conjugation also fixes $x_{-\alpha-\beta}(Xf)$ and transforms $x_{\beta-\alpha}(Xf)$ into a product of generators of type 4, Z.

Now pick a root $\gamma \in \Phi \setminus \Psi$ such that $\alpha + \gamma \in \Phi$.
From~\eqref{eq:Zconj3} in which $\gamma$ is substituted for $\beta$ it follows that  $z_{-\alpha}(X^2f, aX^{-1})$ is a product of generators of type 1, Z, 1, 2, 1, 1, Z, Z\@.
\end{proof}

>>>>

For the rest of this subsection $\Phi$ is a simply-laced root system of rank $\geq 3$ and $\alpha$ is a fixed root of $\Phi$. Notice that in the simply-laced case the assumption $\alpha+\beta\not\in \Phi$ in the definition of $Z_0(\alpha)$ is superfluous, i.\,e. $Z_0(\alpha) = \{ \beta\in\Phi \mid \alpha \perp \beta \}$
(cf.~\cite[Proposition~5.7]{St71}).

@@

We need one more technical definition. We denote by $Z_{\alpha, \beta}$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by elements $x_\gamma(f)$, where $\gamma \in Z_+(\alpha) \setminus \{ \alpha - \beta \}$ and $f \in M[X]$. To simplify notation we call the generators of the subgroup $Z_{\alpha,\beta}$ ``generators of type Z``.

\begin{lemma} \label{image-K-a-b} The image of $K(\alpha, \beta)$ under conjugation by $x_\alpha(aX^{-1})$ is contained in the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by $K(\alpha, \beta)$ and $Z_{\alpha, \beta}$. \end{lemma}
\begin{proof} We need to verify that the conjugate by $x_\alpha(aX^{-1})$ of every generator from~\cref{def:Kab} lies in the specified subgroup.

Let $g = z_\gamma(Xf, \xi)$ be a generator of type $1$ for some $\gamma \in \Phi \setminus \Psi$.
In the case $\gamma \perp \alpha$ this generator commutes with $x_\alpha(aX^{-1})$ by~\cref{Zrels}\eqref{Z4}.
If $\gamma$ is such that $\alpha + \gamma \in \Phi$ we obtain from~\cref{Zrels}\eqref{Z2} that
\begin{equation} z_{\gamma}(Xf, \xi) ^ {x_{\alpha}(aX^{-1})} = x_{\alpha} (- af\xi) \cdot x_{\alpha+\gamma} (N_{\gamma, \alpha}\cdot af) \cdot z_{\gamma}(Xf, \xi). \end{equation}
If, on the other hand, $\gamma$ is such that $\alpha - \gamma \in \Phi$, we obtain from~\cref{Zrels}\eqref{Z3} that
\begin{equation} z_{\gamma}(Xf, \xi) ^ {x_{\alpha}(aX^{-1})} = x_{\alpha} (af\xi) \cdot x_{\alpha-\gamma} (N_{\gamma,-\alpha}\cdot af\xi^2) \cdot z_{\gamma}(Xf, \xi). \end{equation}
By the choice of $\gamma$ both $\alpha + \gamma$ and $\alpha - \gamma$ lie in $Z_+(\alpha) \setminus \{\alpha - \beta\}$, therefore in both cases the expressions in the right-hand side are products of generators of type Z, Z, 1.

A similar computation using \eqref{C1}--\eqref{C3} of \cref{Crels} shows that the conjugate of a generator of type 5 is either the generator itself or a product of generators of type Z, Z, 5.

Now let us verify the assertion for the generators of types 2--4.
By~\eqref{Chevalley-CCF2} the conjugation by $x_\alpha(aX^{-1})$ fixes root subgroups $X_{\alpha}(X\cdot M[X])$, $X_{\beta}(X\cdot M[X])$, $X_{\alpha-\beta}(X\cdot M[X])$. 
On the other hand, from~\eqref{Chevalley-CCF1} we get that
\begin{align*}
 x_{-\beta}(X^2f)^{x_\alpha(aX^{-1})}     &= x_{-\beta}(X^2f) \cdot x_{\alpha-\beta}(N_{-\beta, \alpha} \cdot aXf);\\
 x_{\beta-\alpha}(Xf)^{x_\alpha(aX^{-1})} &= x_{\beta-\alpha}(Xf) \cdot x_{\beta}(N_{\alpha,-\beta}\cdot af).
\end{align*}
Clearly, the expressions in the right-hand side are products of generators of type 2, 4 and 4, Z.

Now pick a root $\gamma \in \Phi \setminus \Psi$ such that $\alpha + \gamma \in \Phi$. Substituting $s = Xf$, $\eta = X$, $\xi = aX^{-1}$ and $\alpha = -\alpha - \gamma$ into~\eqref{Z5} we obtain that
\begin{multline*} z_{-\alpha}(X^2f, aX^{-1}) = x_{-\alpha-\gamma}(\epsilon Xf) \cdot x_{-\gamma}(-af) \cdot x_{\gamma}(aX^2 f) \cdot x_{-\alpha}(X^2f) \cdot \\
 \cdot z_{-\alpha-\gamma}(-\epsilon Xf, -\epsilon a) \cdot x_{\alpha+\gamma}(-\epsilon a^2 Xf) \cdot x_{\alpha}(- a^2 f) \cdot z_{-\gamma}(a f, -X), \end{multline*}
where $\epsilon = N_{\gamma,\alpha}$. 
It is clear that the the first 7 factors in the right-hand side are generators of type
1, Z, 1, 2, 1, 1, Z.

From~\eqref{eq:Zconj3} in which $\gamma$ is substituted for $\beta$ it follows that  $z_{-\alpha}(X^2f, aX^{-1})$ is a product of generators of type 1, Z, 1, 2, 1, 1, Z, Z\@.
\end{proof}

\begin{lemma} \label{P0_conj-alternative} The subgroup $P_\alpha(0)$ is stable under conjugation by $x_\alpha(aX^{-1})$ for arbitrary $a \in A$. \end{lemma}
\begin{proof} It follows from~\cref{K-a-b} that $\widetilde{P}_\alpha(0) \cap K(A[X], M[X]) =  \widetilde{K}(\alpha, \beta) \cdot X_{-\beta}(M \cdot X).$ 
Since any subgroup $H \leq \St(\Phi, A[X], M[X])$ satisfies $H = \left(H \cap K(A[X], M[X])\right) \rtimes ev^*_{X=0}\left(H\right)$, we obtain from~\cref{rem:DS} that
$\widetilde{P}_\alpha(0) =  \left(\widetilde{K}(\alpha, \beta) \cdot X_{-\beta}(M \cdot X)\right) \rtimes Z_\alpha(A, M).$

Thus, in view of~\cref{image-K-a-b} it remains to show that $X_{-\beta}(MX)^{x_\alpha(aX^{-1})} \subseteq P_\alpha(0)$ and $Z_{\alpha, \beta}^{x_\alpha(aX^{-1})} \subseteq P_\alpha(0)$.
\end{proof}
